State of the Practice for Lattice Boltzmann Method Software

### 2021-12-12 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Finished revising up to and including Section 3 (Quantitative Findings)

Sixteen software packages require less than 10 dependencies to be installed. All
but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Eighteen of the software packages do not explicitly indicate software dependency
versions. Some software package installation issues, specifically those
occurring when manual installation of dependencies is required, may be avoided
if versions of dependencies are specified. Fifteen software packages do not have
detailed instructions for installing dependencies.

Sixteen software packages have less than 10 manual installation steps. If
dependencies are installed in one command then none of the software packages
take more than 20 steps to install. The average number of steps is about eight,
and the fewest is two (LB3D-Prime).

All but six (ESPResSo, HemeLB, laboetie, LB3D-Prime, lbmpy, waLBerla) of the
software packages have a way to specifically verify the installation. Most have
some sort of tutorial examples that can be run by the user. Some other ways of
installation validation include validation scripts (LB2D-Prime, lettuce, Ludwig,
LUMA), automatic validation after the installation (LatBo.jl), and instructions
to manually review the file system (LIMBES).

Uninstallation instructions were found for only one of the software packages: pyLBM.

@@

The inclusion of requirements specification and theory documentation greatly
benefits the correctness and verifiability of software packages. The use of
document generation tools can help build confidence in correctness. The addition
of easy to follow tutorials further helps users verify the software and have
confidence in its correctness. Unit testing documentation and capability, as
well as the use of continuous integration tools and techniques such as Bamboo,
Jenkins, and Travis CI, help verify correctness.

@@

The analysis of surface reliability focused on package installation and
tutorials. Errors occurred when installing 16 of the software packages. Every
instance prompted an error message. These messages indicated unrecognized
commands (even when following the installation guide), missing links, missing
dependencies, syntax errors in code files. In some instances the error messages
were vague. Several automatic installation processes could not find and load
dependencies. In these instances the installation tried to access outdated
external repositories. Seven of the installations were recovered and verified,
and one of the installations (LB3D-Prime) was assumed to be recovered due to the
absence of any way to verify it. The installation of eight of the software
packages could not be recovered. Most of these broken installations could not
find external dependencies, encountered system incompatibilities, or displayed
vague error messages.

@@

Overall, lower ranked software packages are lacking clear documentation, testing
or tutorial examples, and descriptive error messages, and have broken
dependencies. Thus, regarding surface reliability, software packages would
benefit from clear up-to-date documentation that specifies all dependencies, the
inclusion of testing and tutorial examples, and the assurance of descriptive
error messages during fault conditions.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages using AHP. Software packages with a high score behaved reasonably in
response to unexpected input as described above. All of the software packages
that installed correctly passed this test. They output descriptive error
messages or did not crash. Software packages with a lower surface robustness
score had not installed correctly, so their robustness score may not be a true
reflection of runtime robustness. Similarly, all software packages that
installed correctly and require plain text input files correctly handled an
unexpected change to these input files, including a replacement of new lines
with carriage returns. Furthermore, nine of the top 10 ranked packages are noted
as being alive. LIMBES is noted as being dead.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other. Due to this, a comparison of
performance is not appropriate. In this project we instead looked through each
software package's artifacts for evidence that performance was considered. The
artifacts of 17 software packages mentioned parallelization. This included GPU
processing and the CUDA parallel computing platform, which were mentioned in the
artifacts of 6 packages (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). GPUs
provide superior processing power and speed compared to CPUs, and are often used
for scientific computing when a large amount of data is involved. The software
package TCLB is implemented in a highly efficient multi-GPU code to achieve
performance suitable for model optimization \citep{rutkowski2020open}. In the
Ludwig package, a so-called mixed mode approach is used where fine-grained
parallelism is implemented on the GPU, and MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. The packages that require
GPU and CUDA have better performance at the expense of installability and
surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 18
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include an expected output.
Most user manuals are in the form of a file that can be downloaded, while some
are rendered on a web-page. Some packages (waLBerla) do not have a user manual,
but do have useful documentation distributed on their web-pages. Expected user
characteristics are documented in four software packages (laboetie, LIMBES,
Ludwig, Palabos). Users are typically scientists or engineers. Their background
is often physics, chemistry, biophysics, or mathematics. All but one of the
packages (LIMBES) have a user support model, and many of them have multiple
avenues of user support. The most popular avenue of support is Git, followed by
email and forums. One software package (OpenLB) has an FAQ page.

@@

Software packages were reviewed for the presence of artifacts. Every type of
artifact or file that is not a code file was recorded. The software packages
were also reviewed for software release and documentation version numbers. This
information could be used to better troubleshoot issues and organize
documentation. All but three software packages (LatBo.jl, LB3D-Prime, MechSys)
have source code release and documentation version numbers.

Information on how code is reviewed, or how to contribute to the project was
also noted. In total, 11 software packages have this information, which was
found in various artifacts, including in developer guides, contributor guides,
user guides, developer web-pages, and README files.

@@

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}.
Packages with a higher percentage of comments were designated as more
maintainable. Comments represent more than 10 percent of code files in 15
packages, and the average percentage of code comments is about 14 percent. Four
of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, OpenLB) have
more than the average. Fifth ranked LUMA has only 0.2 percent comments, the
fewest of any package. The package has the most lines of source code, with over
four million. The next largest package is ESPResSo++ with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages using AHP. Software packages with a high score provide version
numbers on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to reasonably handle tracked issues, having most of their
issues closed. Their code files are well commented with more than 10 percent of
the code being comments. Furthermore, nine of the top 10 ranked packages are
noted as being alive. MP-LABS is noted as being dead.\\

\begin{table}
\begin{center}
	\begin{tabular}{ p{3.5cm}p{3.5cm}p{3.5cm}p{2.5cm} }
		\hline
		Name & $\%$ Issues Closed & $\%$ Code Comments& Status\\
		\hline
		DL\_MESO (LBE) & Not Git & 8.06&Alive\\
		ESPResSo & 89.26 & 21.78&Alive\\
		ESPResSo++ & 66.28 & 17.10&Alive\\
		HemeLB & No Issues & 16.68&Dead\\
		laboetie & 18.75 & 2.47&Dead\\		
		LatBo.jl & 93.33 & 0.40&Dead\\
		LB2D-Prime & Not Git & 13.61&Dead\\
		LB3D & Not Git & 13.76&Dead\\
		LB3D-Prime & Not Git & 14.34&Dead\\
		lbmpy& 58.33  & 2.03  &Alive\\
		lettuce & 33.33 & 8.19&Alive\\
		LIMBES & Not Git & 17.39&Dead\\
		Ludwig& 60.00 & 20.70&Alive\\
		LUMA& 85.71   & 0.20&Alive\\
		MechSys & Not Git & 15.11&Alive\\
		MP-LABS & 100.00 & 26.67&Dead\\
		OpenLB & Not Git & 22.43&Alive\\
		Palabos & 89.47 & 17.76&Alive\\
		pyLBM & 66.67& 16.12&Alive\\
		Sailfish & 22.22 & 9.26&Alive\\
		SunlightLB & Not Git & 17.67&Dead\\
		TCLB & 60.32 & 6.02&Alive\\
		waLBerla & 72.90 & 22.62&Alive\\
		\hline
	\end{tabular}
	\caption{Git Repository Data} \label{gitrepodata}
\end{center}
\end{table}

@@

Each software package was measured for the total number of source code files. A
larger number of source files was associated with increased reusability due to
increased modularization. Some packages have more features than others,
consequently contributing to reusablility since they have more source code that
can be reused. The software packages were also reviewed for the presence of API
documentation, which indicates that a software package was developed with
interaction between other software applications in mind.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages using AHP. Software packages with a high score have thousands of source
code files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations as
well as modeling that does not use LBM. For this reason a comparison with other
software packages is not on a level field. However, these packages do have an
abundance of reusable components. Furthermore, nine of the top 10 ranked
packages are noted as being alive. HemeLB is noted as being dead.

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\hline
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\hline
			DL\_MESO (LBE) & 310 & 51 & 170223& 549\\
			ESPResSo &1309& 86 & 186700&143\\
			ESPResSo++ &5328& 66 & 969196&182\\
			HemeLB &1065& 48 & 95104&89\\
			laboetie &133& 1 & 48403&364\\		
			LatBo.jl &41& 0 & 42172&1029\\
			LB2D-Prime &82& 19 & 54755&668\\
			LB3D &99& 76 & 39766&402\\
			LB3D-Prime &23& 6 & 12944&563\\
			lbmpy&201&  28 & 46489  &231\\
			lettuce &62& 0 & 5529&89\\
			LIMBES &26& 1 & 4872&187\\
			Ludwig&859& 32 & 109811&128\\
			LUMA&312& 19 & 4370670&14000\\
			MechSys &324& 3 & 85543&264\\
			MP-LABS &307& 3 & 43124&140\\
			OpenLB &1104& 5 & 209034&189\\
			Palabos &1829& 67 & 547623&299\\
			pyLBM &258& 85& 32314&125\\
			Sailfish &632& 11 & 69398&110\\
			SunlightLB & 36& 1 & 7646& 212\\
			TCLB &535& 7 & 43226&81\\
			waLBerla & 2395 & 67 & 848146&353\\
			\hline
		\end{tabular}
		\caption{Module Data} \label{moduledata}
	\end{center}
\end{table}

Table~\ref{moduledata} shows file and line of code data of the software
packages. Packages with a high reusability score do not have many LOC per text
file, generally having a few hundred lines or less. This suggests that the
source code of these packages is likely functionally modularized, and modules
could be reused in other projects.

There was a strong focus on modularity when designing the waLBerla framework to
enhance productivity, reusability, and maintainability
\citep{bauer2021walberla}. Its software design has enabled waLBerla to be
successfully applied in several projects as a basis for various extensions
\citep{bauer2021walberla}.

@@

Ten random source code files of each software package were reviewed for several
measures. This assessment of surface understandability may not perfectly reflect
each package due to the practical limitation of only being able to test 10
files.

All of the packages appear to have a consistent indentation and formatting
style. Only LUMA and HemeLB explicitly identify coding standards that are used
during development. Generally, the software packages use consistent,
distinctive, and meaningful code identifiers. Only four packages (LB2D-Prime,
LB3D-Prime, LIMBES, MP-LABS) appear to use vague identifiers, such as single
letters for variables. Symbolic constants were observed in the source code of 12
packages. The constants are used for various parameters, mathematical constants,
and matrix definitions. All of the packages are well commented and the comments
clearly indicate what is being done. Domain algorithms are noted in the source
code of 11 packages. Table~\ref{moduledata} suggests that the software packages
are modularized to various degrees. When observing the source code files, it was
found that 13 of the packages have a consistent style and order of function
parameters.

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking of the software packages
using AHP. Software packages with an overall high score had ranked high in at
least several of the individual qualities that were quantitatively measured. The
overall ranking is found with an assumption of an equal weighting of all
qualities. If the qualities were weighed differently, the overall software
package ranking would change.

Looking at the top three ranked packages, Ludwig scored high in surface
correctness and verifiability, surface robustness, surface usability,
maintainability, and visibility and transparency. ESPResSo had achieved a
relative high score in installability, surface correctness and verifiability,
surface usability, maintainability, reusability, and visibility and
transparency. Palabos scored high in installability, surface reliability,
surface robustness, maintainability, and understandability.

Section~\ref{qualityrecommentations} further analyzes these findings and offers
some software quality recommendations for future development of LBM software.

>>>>

Sixteen software packages require less than 10 dependencies to be installed. All
but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Eighteen of the software packages do not explicitly indicate software dependency
versions. Some software package installation issues, specifically those
occurring when manual installation of dependencies is required, may be avoided
if versions of dependencies are specified. Fifteen software packages do not have
detailed instructions for installing dependencies. Sixteen software packages
have less than 10 manual installation steps. If dependencies are installed in
one command then none of the software packages take more than 20 steps to
install. The average number of steps is about eight, and the fewest is two
(LB3D-Prime).

All but six (ESPResSo, HemeLB, laboetie, LB3D-Prime, lbmpy, waLBerla) of the
software packages have a way to verify the installation. Most have some sort of
tutorial examples that can be run by the user. Some other ways of installation
validation include validation scripts (LB2D-Prime, lettuce, Ludwig, LUMA),
automatic validation after the installation (LatBo.jl), and instructions to
manually review the file system (LIMBES).  Uninstallation instructions were
found for only one of the software packages: pyLBM.

@@

The inclusion of requirements specification and theory documentation greatly
benefits the correctness and verifiability of software packages. The use of
document generation tools can help build confidence in correctness. The addition
of easy to follow tutorials further helps users verify the software and have
confidence in its correctness. Unit testing, as well as the use of continuous
integration tools and techniques such as Bamboo, Jenkins, and Travis CI, help
verify correctness.

@@

The analysis of surface reliability focused on package installation and
tutorials. Errors occurred when installing 16 of the software packages. Every
instance prompted an error message. These messages indicated unrecognized
commands (even when following the installation guide), missing links, missing
dependencies, and syntax errors in code files. In some instances the error messages
were vague. Several automatic installation processes could not find and load
dependencies. In these instances the installation tried to access outdated
external repositories. Seven of the installations were recovered and verified,
and one of the installations (LB3D-Prime) was assumed to be recovered due to the
absence of any way to verify it. The installation of eight of the software
packages could not be recovered. Most of these broken installations could not
find external dependencies, encountered system incompatibilities, or displayed
vague error messages.

@@

Overall, lower ranked software packages are lacking clear documentation, testing
or tutorial examples, and descriptive error messages, and have broken
dependencies. Thus, regarding surface reliability, software packages would
benefit from clear up-to-date documentation that specifies all dependencies, the
inclusion of testing and tutorial examples, and the assurance of descriptive
error messages after faults.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages using AHP. Software packages with a high score behaved reasonably in
response to unexpected input as described above. All of the software packages
that installed correctly passed this test. They output descriptive error
messages or did not crash. Software packages with a lower surface robustness
score did not install correctly, so their robustness score may not be a true
reflection of runtime robustness. All successfully installed software packages
that require a plain text input file correctly handled an unexpected change to
these input files, including a replacement of new lines with carriage returns.
Furthermore, nine of the top 10 ranked packages are noted as being alive. LIMBES
is noted as being classified as dead.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of performance is not
appropriate. In this project we instead looked through each software package's
artifacts for evidence that performance was considered. The artifacts of 17
software packages mentioned parallelization. This included GPU processing and
the CUDA parallel computing platform, which were mentioned in the artifacts of 6
packages (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). GPUs provide
superior processing power and speed compared to CPUs, and are often used for
scientific computing when a large amount of data is involved. The software
package TCLB is implemented in a highly efficient multi-GPU code to achieve
performance suitable for model optimization \citep{rutkowski2020open}. In the
Ludwig package, a so-called mixed mode approach is used where fine-grained
parallelism is implemented on the GPU, and MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. The packages that require
GPU and CUDA have better performance at the expense of installability and
surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 18
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output.
Most user manuals are in the form of a file that can be downloaded, while some
are rendered on a web-page. Some packages (waLBerla) do not have a user manual,
but do have useful documentation distributed on their web-pages. Expected user
characteristics are documented in four software packages (laboetie, LIMBES,
Ludwig, Palabos). Users are typically scientists or engineers. Their background
is often physics, chemistry, biophysics, or mathematics. All but one of the
packages (LIMBES) have a user support model, and many of them have multiple
avenues of user support. The most popular avenue of support is Git, followed by
email and forums. One software package (OpenLB) has an FAQ page.

@@

Software packages were reviewed for the presence of artifacts. Every type of
artifact or file that is not a code file was recorded. The software packages
were also reviewed for software release and documentation version numbers. This
information can be used to troubleshoot issues and organize documentation. All
but three software packages (LatBo.jl, LB3D-Prime, MechSys) have source code
release and documentation version numbers.

When present, information on how code is reviewed, or how to contribute to the
project was also noted. In total, 11 software packages have this information,
which was found in various artifacts, including in developer guides, contributor
guides, user guides, developer web-pages, and README files.

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{3.5cm}p{3.5cm}p{2.5cm} }
			\hline
			Name & $\%$ Issues Closed & $\%$ Code Comments & Status\\
			\hline
			DL\_MESO (LBE) & Not Git & 8.06& Alive\\
			ESPResSo & 89.26 & 21.78 & Alive\\
			ESPResSo++ & 66.28 & 17.10 & Alive\\
			HemeLB & No Issues & 16.68 & Dead\\
			laboetie & 18.75 & 2.47 & Dead\\		
			LatBo.jl & 93.33 & 0.40 & Dead\\
			LB2D-Prime & Not Git & 13.61 & Dead\\
			LB3D & Not Git & 13.76 & Dead\\
			LB3D-Prime & Not Git & 14.34 & Dead\\
			lbmpy& 58.33  & 2.03 & Alive\\
			lettuce & 33.33 & 8.19 & Alive\\
			LIMBES & Not Git & 17.39 & Dead\\
			Ludwig& 60.00 & 20.70 & Alive\\
			LUMA& 85.71   & 0.20 & Alive\\
			MechSys & Not Git & 15.11 & Alive\\
			MP-LABS & 100.00 & 26.67 & Dead\\
			OpenLB & Not Git & 22.43 & Alive\\
			Palabos & 89.47 & 17.76 & Alive\\
			pyLBM & 66.67& 16.12 & Alive\\
			Sailfish & 22.22 & 9.26 & Alive\\
			SunlightLB & Not Git & 17.67 & Dead\\
			TCLB & 60.32 & 6.02 & Alive\\
			waLBerla & 72.90 & 22.62 & Alive\\
			\hline
		\end{tabular}
		\caption{Git Repository Data} \label{gitrepodata}
	\end{center}
	\end{table}

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}.
Packages with a higher percentage of comments were designated as more
maintainable. Comments represent more than 10 percent of code files in 15
packages, and the average percentage of code comments is about 14 percent. Four
of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, OpenLB) have
more than the average. Fifth ranked LUMA has only 0.2 percent comments, the
fewest of any package. This package has the most lines of source code, with over
four million. The next largest package is ESPResSo++ with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages using AHP. Software packages with a high score provide version
numbers on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to reasonably handle issue tracking, having most of their
issues closed. Their code files are well commented with more than 10 percent of
the code being comments. Furthermore, nine of the top 10 ranked packages are
noted as being alive. MP-LABS is noted as being dead.\\

\wss{Sort ``Git Repository Data'' table in descending order of percentage of
code comments.}

@@

We measured the total number of source code files for each project. A larger
number of source files is associated with increased reusability, due to our
assumption that this indicates increased modularization. Some packages have more
features than others.  This is assumed to contribute to reusablility, since they have
more source code for potential reuse. The software packages were also reviewed
for the presence of API documentation, which indicates that a software package
was developed with interaction between other software applications in mind.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages using AHP. Software packages with a high score have thousands of source
code files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modeling. For this reason, a comparison with other software packages is
not on a level field. However, these packages do have an abundance of reusable
components. Furthermore, nine of the top 10 ranked packages are noted as being
alive. HemeLB is considered to be dead.

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\hline
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\hline
			DL\_MESO (LBE) & 310 & 51 & 170223 & 549\\
			ESPResSo & 1309& 86 & 186700 & 143\\
			ESPResSo++ & 5328& 66 & 969196 & 182\\
			HemeLB & 1065& 48 & 95104 & 89\\
			laboetie & 133& 1 & 48403 & 364\\		
			LatBo.jl & 41& 0 & 42172 & 1029\\
			LB2D-Prime & 82& 19 & 54755 & 668\\
			LB3D & 99 & 76 & 39766 & 402\\
			LB3D-Prime & 2 3& 6 & 12944 & 563\\
			lbmpy& 201 & 28 & 46489 & 231\\
			lettuce & 62 & 0 & 5529 & 89\\
			LIMBES & 26 & 1 & 4872 & 187\\
			Ludwig & 859 & 32 & 109811 & 128\\
			LUMA & 312 & 19 & 4370670 & 14000\\
			MechSys & 324 & 3 & 85543 & 264\\
			MP-LABS & 307 & 3 & 43124 & 140\\
			OpenLB & 1104 & 5 & 209034 & 189\\
			Palabos & 1829 & 67 & 547623 & 299\\
			pyLBM & 258 & 85 & 32314 & 125\\
			Sailfish & 632 & 11 & 69398 & 110\\
			SunlightLB & 36 & 1 & 7646 & 212\\
			TCLB & 535 & 7 & 43226 & 81\\
			waLBerla & 2395 & 67 & 848146 & 353\\
			\hline
		\end{tabular}
		\caption{Module Data} \label{moduledata}
	\end{center}
\end{table}

Table~\ref{moduledata} shows file and Lines Of Code (LOC) data for the software
packages. Packages with a high reusability score do not have as many LOC per
text file, generally having a few hundred lines or less. This suggests that the
source code of these packages is likely functionally modularized, and modules
could be reused in other projects.

There was a strong focus on modularity when designing the waLBerla framework to
enhance productivity, reusability, and maintainability
\citep{bauer2021walberla}. The design of waLBerla has enabled it to be
successfully applied in several projects as a basis for various extensions
\citep{bauer2021walberla}.

@@

Ten random source code files of each software package were reviewed for several
measures. This assessment of surface understandability may not perfectly reflect
each package, due to the practical limitation of only examining 10 files.

All of the packages appear to have consistent indentation and formatting. Only
LUMA and HemeLB explicitly identify coding standards that are used during
development. Generally, the software packages use consistent, distinctive, and
meaningful code identifiers. Only four packages (LB2D-Prime, LB3D-Prime, LIMBES,
MP-LABS) appear to use vague identifiers, such as single letters for variables.
Symbolic constants were observed in the source code of 12 packages. The
constants are used for various parameters, mathematical constants, and matrix
definitions. All of the packages are reasonably well commented, with the
comments clearly indicating what is being done (as opposed to how it is being
done). Domain algorithms are noted in the source code of 11 packages.
Table~\ref{moduledata} suggests that the software packages are modularized to
various degrees. When observing the source code files, it was found that 13 of
the packages have a consistent style and order of function parameters.

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking of the software packages
using AHP. In the absence of other information on priorities, the overall
ranking was calculated by assuming an equal weighting between all qualities. If the
qualities were weighed differently, the overall software package ranking would
change. Software packages with an overall high score ranked high in at least
several of the individual qualities that were quantitatively measured.

Looking at the top three ranked packages: Ludwig scored high in surface
correctness and verifiability, surface robustness, surface usability,
maintainability, and visibility and transparency. ESPResSo had achieved a
relative high score in installability, surface correctness and verifiability,
surface usability, maintainability, reusability, and visibility and
transparency. Palabos scored high in installability, surface reliability,
surface robustness, maintainability, and understandability.
